Under-approximation of reachability in multivalued asynchronous networks

### 2013-04-02 parts/sa.tex
compression of GLC explanations

The reachability of a process $a_i$ is approximated by the ability to solve all objectives $\PHobjp{a}{j}{i}$ so that $a_j$ is in the initial context.
To this end, the abstractions of bounce sequences (\pref{def:aBS}) solving these objectives are computed, leading to new required processes, and thus new objectives on other sorts.
Thus, the graph of local causality $\cwB = (\V, \E)$ defined in \pref{def:glc} is a graph where $\V \subset \Proc \cup \Obj \cup \Sol$ and $\E \subset \V \times \V$
in which edges link any required process $a_i$ in $\Proc$ to all corresponding objectives $P$ in $\Obj$,
and any objective $P$ to abstractions of bounce sequences $ps \in \Sol$ (\pref{def:aBS}) solving this objective;
such sets $ps$ then stand for new required processes.
In the case where a process $a_i$ belongs to a cooperative sort, it is directly linked to the set of the processes $\csState(a_i)$ (\pref{def:csState}) that it represents.
If the solving of an objective $P$ requires a process of $\PHsort(P)$, a continuity edge is added to re-target this objective (\pref{def:maxCont}).
Finally, as the active process of every sort may evolve, $\cwB$ is obtained iteratively by saturation with every process it contains,
\ie by overriding its initial context $\ctx$ by $\allprocs(\V, \E)$, defined by:
$$\allprocs(\V, \E) = (V \cap \Proc) \cup \{ \PHtarget(P), \PHbounce(P) \mid P \in \V \cap \Obj \}$$

@@

\pref{eq:Vw}, \eqref{eq:Vproc} and \eqref{eq:VE} ensure that all required nodes are in $\cwV$.
Elements in $\PHproc$ are required processes,
elements in $\Obj$ are the objectives to reach these processes
and elements in $\Sol$ are the sets of processes that allow to solve these objectives.
\pref{eq:ESol1} links an objective $P$ to its solutions in $\BS(P)$
and \pref{eq:ESol2} links a solution $ps$ to the processes $a_i \in ps$ it contains.
Then, \pref{eq:EReq} links a required component process $a_i$ to all objectives $\PHobjp{a}{j}{i}$ with $a_j \in \ctx$
while \pref{eq:EPrio} links a required cooperative sort process $a_i$ to its solutions in $\csState(a_i)$.
Finally, if the resolution of an objective $P$ requires a process of $\PHsort(P)$, then \pref{eq:ECont} re-targets this objective ($\gCont$).

@@

\begin{example}
  Let $\PH' = (\PHs, \PHl, \PHh'^{\langle 1 \rangle})$ be the “flattened” version of the PH in \pref{fig:ph-livelock}, that is: $\PHh'^{\langle 1 \rangle} = \PHh^{(1)} \cup \PHh^{(2)}$,
  which is equivalent to a PH in the semantics without priorities.
  The under-approximation on such models developed in~\cite{PMR12-MSCS} concludes that $c_1$ is reachable from $\ctx = \PHstate{a_1, b_0, c_0, ab_{10}}$;
  this is due to an over-approximation of the behaviours inherent to cooperative sorts in the semantics without priorities.

  This unwanted behaviour is palliated by the semantics of PH with priorities proposed in this paper.
  Indeed, the under-approximation given in \pref{th:approxinf} does not conclude regarding the reachability of $c_1$,
  as one of the edges of the resulting graph of local causality is not coherent (\pref{def:coherent}),
  as shown in in \pref{fig:sa-livelock}.
  (However, from the inconclusiveness of \pref{th:approxinf}, one cannot conclude about the unreachability of $c_1$.
  Such analysis should be driven for instance with over-approximation methods developed in~\cite{PMR10-TCSB}.)

  We note that if the actions $\PHhit{a_0}{b_0}{b_1}$ and $\PHhit{b_0}{a_0}{a_1}$ are replaced by the self-hits
  $\PHhit{a_0}{a_0}{a_1}$ and $\PHhit{b_0}{b_0}{b_1}$,
  then \pref{th:approxinf} concludes that $c_1$ is reachable from $\ctx$.
\end{example}

>>>>

The graph of local causality $\cwB = (\V, \E)$ defined in \pref{def:glc} is a graph where $\V \subset \Proc \cup \Obj \cup \Sol$ and $\E \subset \V \times \V$.
A node in $\PHproc$ is a required process,
a node in $\Obj$ is an objective to reach a given process
and a node in $\Sol$ is a sets of processes required for the solving.
An objective $P \in \Obj$ is solvable if the abstractions of bounce sequences $\aBS(P) \in \Sol$ (\pref{def:aBS}) can be reached (\pref{eq:ESol1}), leading to several required processes (\pref{eq:ESol2}).
If $a$ is a component, the reachability of a process $a_i$ is approximated by the ability to solve all objectives $\PHobjp{a}{j}{i} \in \Obj$ for all $a_j$ in the initial context (\pref{eq:EReq});
if $a$ is a cooperative sort, the reachability of $a_i$ is directly solved by the set of the processes $\csState(a_i)$ (\pref{def:csState}) it represents (\pref{eq:EPrio}).
Finally, if the resolution of an objective $P$ requires a process of $\PHsort(P)$, \ie $\gCont(\PHsort(P), P) \neq \emptyset$ (\pref{def:maxCont}), then this objective is re-targeted (\pref{eq:ECont}).
\pref{eq:Vw}, \eqref{eq:Vproc} and \eqref{eq:VE} ensure that all required nodes are in $\cwV$.

\begin{comment}
The reachability of a process $a_i$ is approximated by the ability to solve all objectives $\PHobjp{a}{j}{i}$ for all $a_j$ in the initial context.
To this end, the abstractions of bounce sequences (\pref{def:aBS}) solving these objectives are computed, leading to new required processes, and thus new objectives on other sorts.
Thus, the graph of local causality $\cwB = (\V, \E)$ defined in \pref{def:glc} is a graph where $\V \subset \Proc \cup \Obj \cup \Sol$ and $\E \subset \V \times \V$
in which edges link any required process $a_i$ in $\Proc$ to all corresponding objectives $P$ in $\Obj$,
and any objective $P$ to abstractions of bounce sequences $ps \in \Sol$ (\pref{def:aBS}) solving this objective;
such sets $ps$ then stand for new required processes.
In the case where a process $a_i$ belongs to a cooperative sort, it is directly linked to the set of the processes $\csState(a_i)$ (\pref{def:csState}) that it represents.
If the solving of an objective $P$ requires a process of $\PHsort(P)$, a continuity edge is added to re-target this objective (\pref{def:maxCont}).
Finally, as the active process of every sort may evolve, $\cwB$ is obtained iteratively by saturation with every process it contains,
\ie by overriding its initial context $\ctx$ by $\allprocs(\V, \E)$, defined by:
$$\allprocs(\V, \E) = (V \cap \Proc) \cup \{ \PHtarget(P), \PHbounce(P) \mid P \in \V \cap \Obj \}$$
\end{comment}

@@

\begin{comment}
\pref{eq:Vw}, \eqref{eq:Vproc} and \eqref{eq:VE} ensure that all required nodes are in $\cwV$.
Elements in $\PHproc$ are required processes,
elements in $\Obj$ are the objectives to reach these processes
and elements in $\Sol$ are the sets of processes that allow to solve these objectives.
\pref{eq:ESol1} links an objective $P$ to its solutions in $\BS(P)$
and \pref{eq:ESol2} links a solution $ps$ to the processes $a_i \in ps$ it contains.
Then, \pref{eq:EReq} links a required component process $a_i$ to all objectives $\PHobjp{a}{j}{i}$ with $a_j \in \ctx$
while \pref{eq:EPrio} links a required cooperative sort process $a_i$ to its solutions in $\csState(a_i)$.
Finally, if the resolution of an objective $P$ requires a process of $\PHsort(P)$, then \pref{eq:ECont} re-targets this objective ($\gCont$).
\end{comment}

@@

\begin{example}
  Let $\PH' = (\PHs, \PHl, \PHh'^{\langle 1 \rangle})$ be the “flattened” version of the PH in \pref{fig:ph-livelock}, that is: $\PHh'^{\langle 1 \rangle} = \PHh^{(1)} \cup \PHh^{(2)}$,
  which is equivalent to a PH in the semantics without priorities.
  Due to spurious behaviours inherent to the cooperative sorts in this semantics, 
  the original under-approximation developed in~\cite{PMR12-MSCS} concludes that $c_1$ is reachable in $\PH'$ from $\ctx = \PHstate{a_1, b_0, c_0, ab_{10}}$.
  %this is due to an over-approximation of the behaviours inherent to cooperative sorts in the semantics without priorities.

  Such unwanted behaviours are palliated by the semantics of PH with priorities proposed in this paper.
  Indeed, the under-approximation given in \pref{th:approxinf} does not conclude regarding the reachability of $c_1$,
  as one of the edges of the resulting graph of local causality is not coherent (\pref{def:coherent}),
  as shown in in \pref{fig:sa-livelock}.
  (However, from the inconclusiveness of \pref{th:approxinf}, one cannot conclude about the unreachability of $c_1$.
  Such analysis should be driven for instance with over-approximation methods developed in~\cite{PMR12-MSCS}.)

  However, if $\PHhit{a_0}{b_0}{b_1}$ and $\PHhit{b_0}{a_0}{a_1}$ are replaced by
  $\PHhit{a_0}{a_0}{a_1}$ and $\PHhit{b_0}{b_0}{b_1}$,
  then \pref{th:approxinf} concludes that $c_1$ is reachable from $\ctx$.
\end{example}
